---
title: "Project2"
author: "Joyce Aldrich"
date: "2022-10-09"
output:
  html_document: default
  pdf_document: default
---

# example 1. loading csv into r
```{r}
library(readr)
city_weekly_temperature_wide <-read_csv(file = '/Users/joycealdrich/Documents/SPS Data Science/Data 607/Project_2/city_week_temp.csv')
head(city_weekly_temperature_wide)
```

# reshaping data to long format
```{r}
library(tidyr)
city_weekly_temperature_long <- city_weekly_temperature_wide %>%
gather(day,temperature, -c(city))
head(city_weekly_temperature_long)

```
# adding the average temperature for each city 
```{r}
library(dplyr)
city_weekly_temperature_long <- city_weekly_temperature_long %>%
  group_by(city) %>%
  mutate(average_temperature = mean(temperature),
         temperature_diff = temperature - mean(temperature)
         )
head(city_weekly_temperature_long)
```

# using ggplot geom_point to see each city's weekly temperature
```{r}
library(ggplot2)
ggplot(data= city_weekly_temperature_long ) +
  geom_point(mapping = aes(x=day, y=temperature, color=city)) +
  facet_wrap(~city, nrow=1)
```

# compare 6 cities' weekly temperature and place the overall mean in the graph  
```{r}
library(tidyverse)
library(ggplot2)

city_weekly_temperature_long %>%
  ggplot(aes(x=day, y=temperature,group=city,colour=city))+
  geom_point(size =3, alpha=0.3)+
  geom_line(size =0.5)+
  geom_hline(aes(yintercept=mean(temperature)))+
  theme_minimal()+
  labs(title ="Weekly Temperature of Cities")
  
```

# example 2: 
# loading cvs file into r
```{r}

library(readr)
exam_results <-read_csv(file = '/Users/joycealdrich/Documents/SPS Data Science/Data 607/Project_2/exam_result.csv')
head(exam_results)
```

# reshaping to the long format
```{r}
library(tidyr)
exam_results_long<- exam_results %>%
  gather(exam,score, -c(Student,Gender))
head(exam_results_long)
```
# adding score_average_by_student col & removing the na row
```{r}
exam_results_long <- exam_results_long %>%
  group_by(Student) %>%
  filter(!is.na(score)) %>%
  mutate(Score_average_by_student = mean(score),
         )
head(exam_results_long)
```
# adding max_socre, min_score ,core_average_by_exam col 
```{r}
exam_results_long <- exam_results_long %>%
  group_by(exam) %>%
  filter(!is.na(score)) %>%
  mutate(Max_score_by_exam = max(score),
         Min_score_by_exam = min(score),
         average_score_by_exam = mean (score)
         )
head(exam_results_long)
  
```
# compare each exam score by each student and place an overall mean in the graph  
```{r}
library(tidyverse)
library(ggplot2)

exam_results_long %>%
  ggplot(aes(x=exam, y=score, group=Student, colour=Student))+
  geom_point(size =3, alpha=0.3)+
  geom_hline(aes(yintercept=mean(score)))+
  theme_minimal()+
  labs(title ="score of three exams")
  
```

# example 3:
# loading excel(format download) file into r
```{r}
library(readxl)
gdp_by_county <- read_excel("gdp_by_county.xlsx")
head(gdp_by_county)

```

# updating the col name and remove NA rows
```{r}
names(gdp_by_county)[6]<-paste("2012")
names(gdp_by_county)[7]<-paste("2013")
names(gdp_by_county)[8]<-paste("2014")
names(gdp_by_county)[9]<-paste("2015")

gdp_by_county <- gdp_by_county[-c(1,2),]
head(gdp_by_county)
```
# keeping consolidation GDP amount
```{r}
gdp_by_county_wide<- gdp_by_county %>%
filter(LineCode == 1) 
head(gdp_by_county_wide)
```

# reshaping to long data format
```{r}
library(tidyr)
gdp_by_county_long <- gdp_by_county_wide %>%
  gather(year, amount, -c(1:5))
head(gdp_by_county_long)
```
# random select 6 sample counties in AL state 
```{r}
gdp_by_county_long_1 <- gdp_by_county_long %>%
  filter(Postal =="AL" & FIPS %in% c('01005','01011','01017','01019','01061','01063'))
```
# change class for amount var
```{r}

gdp_by_county_long_1 <-gdp_by_county_long_1 %>%
   mutate(amount=as.integer(amount))
```
# insert the col for growth rate
```{r}
library(dbplyr)
gdp_by_county_long_1 <- gdp_by_county_long_1 %>%
  group_by(Countyname)%>%
  mutate(Growth=(amount-lag(amount))/lag(amount))
head(gdp_by_county_long_1)
```


# creating ggplot 
```{r}
library(tidyverse)
library(ggplot2)

gdp_by_county_long_1 %>%
  ggplot(aes(x=year, y=amount,color=Countyname))+
  geom_point(size =5, alpha=1)+
  geom_hline(aes(yintercept=mean(amount)))+
  theme_minimal()+
  labs(title ="Alabama- 6 Counties - 2012-2015 GDP")
```

